Estimating demographic bias on tests of children's early vocabulary

### 2022-01-16 paper/cogsci.Rmd
writing

The outline of this paper is as follows.
First, we will introduce the Wordbank data and the IRT model we use to analyze the data.
We then fit the IRT model to each group along each dimension, and examine the parameters for evidence of DIF

@@

>>>>

The outline of this paper is as follows.
First, we will introduce the Wordbank data and the IRT model we use to analyze the data.
We then fit the IRT model to each group along each demographic dimension, and examine the parameters for evidence of DIF, noting in particular how many items are significantly biased in favor of each group.
We identify a set of suspect items to eliminate (or in the future, replace), and provide updated estimates of the effect size of these demographic variables, were the biased items to be eliminated.
Next, using a corpus of child-directed speech, we examine the extent to which variation in word frequencies heard by boys vs. girls predicts sex-based DIF.
Finally, we provide recommendations for next steps to be taken to identify and replace biased items on the CDI.

@@

```{r load-data}
load("../data/glimmer_prodWS_models.Rds")

library(mirt)

mod_intuitive_eth
mod_intuitive_ses
mod_intuitive_sex
```

```{r}
# baseline Rasch model differences in ability by demographic groups
```

```{r}
mod_intuitive_ses <- fit_mod_intuitive(d_mat, ses_group)
```

```{r demo-ability}
# show density plots of child ability based on group models
# sex
ses <- d_demo %>% filter(!is.na(sex_group), production!=0) %>%
  mutate(theta_sex = fscores(mod_intuitive_sex)) %>%
  rename(`Sex Group`=sex_group)
# cor.test(eth$production, eth$theta_eth) # .98

sex_m <- sex %>% group_by(`Sex Group`) %>%
  summarise(ability = mean(theta_sex, na.rm=T))

p1 <- sex %>%
  ggplot(aes(x=theta_sex, color=`Sex Group`, fill=`Sex Group`)) +
    geom_density(alpha=.2) + theme_classic() + xlab("Ability") + 
    geom_vline(data=sex_m, aes(xintercept=ability, color=`Sex Group`))

# SES
ses <- d_demo %>% filter(!is.na(ses_group), production!=0) %>%
  mutate(theta_ses = fscores(mod_intuitive_ses)) %>%
  rename(`SES Group`=ses_group)
# cor.test(eth$production, eth$theta_eth) # .98

ses_m <- ses %>% group_by(`Ethnic Group`) %>%
  summarise(ability = mean(theta_ses, na.rm=T))

p2 <- ses %>%
  ggplot(aes(x=theta_ses, color=`SES Group`, fill=`SES Group`)) +
    geom_density(alpha=.2) + theme_classic() + xlab("Ability") + 
    geom_vline(data=ses_m, aes(xintercept=ability, color=`SES Group`))

# ethnicity
eth <- d_demo %>% filter(!is.na(eth_group), production!=0) %>%
  mutate(theta_eth = fscores(mod_intuitive_eth)) %>%
  rename(`Ethnic Group`=eth_group)
# cor.test(eth$production, eth$theta_eth) # .98

eth_m <- eth %>% group_by(`Ethnic Group`) %>%
  summarise(ability = mean(theta_eth, na.rm=T))

p3 <- eth %>%
  ggplot(aes(x=theta_eth, color=`Ethnic Group`, fill=`Ethnic Group`)) +
    geom_density(alpha=.2) + theme_classic() + xlab("Ability") + 
    geom_vline(data=eth_m, aes(xintercept=ability, color=`Ethnic Group`))

ggarrange(p1, p2, p3)
```
